---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

```{r, include=FALSE}
library(classdata)

```

## Step 1 Results

**What variables are there?**

Parcel ID, Address, Style, Occupancy, Sale Date, Sale Price, Multi-sale, Year Built, Acres, Total Living Area (SF), Bedrooms, Finished BSMT (SQ), Lot Area (SF), AC, Fireplace, Neighborhood

**What type are the variables?**

There is a mix of categorical (ex. Neighborhood, Occupancy, Style) and numerical (ex. Total Living Area (SF), Bedrooms, Finished BSMT (SQ), Lot Area (SF)). Some of the variables are continuous.

**What does each variable mean? What do we expect their data range to be?**

Parcel ID: The house ID number in the system, it is a categorical variable and there is no range.

Address: The address the house is located at, it is a categorical variable and there is no range.

Style: The architectural style the house was built in, it is a categorical variable, and the range is 12 (there are 12 different styles the house can be classified as)

Occupancy: How many people live in the house, it is a categorical variable, and the range is 5

Sale Date: The date the house was sold, it is a categorical variable, and the range is any date between July 3rd, 2017 and August 31st, 2022.

Sale Price: The price the building was sold for, it is a numerical variable, and the range is 0 - 22 million

Multi-sale: States if the house has been sold multiple times, it is a numerical variable, and the range is 0 (no), 1 (yes).

Year Built: The year the building was built, it is a numerical variable, and the range is 1880 to 2022 (where 0 is present as a missing value).

Acres: The total acres on the property, it is a numerical variable, and the range is 0 to 12.

Total Living Area (SF): The total living area square footage, it is a numerical variable, and the range is 0 to 6,007.

Bedrooms: The number of bedrooms in the property, it is a numerical variable, and the range is 0 to 10.

Finished BSMT (SQ): The finished basement square footage, it is a numerical variable, and the range is 10 to 6,496 (with NA for non-finished basements).

Lot Area (SF): The square footage of the lot the house was built on, it is a numerical variable, and the range is 0 to 523,228

AC: States whether the house has air conditioning, it is a numerical variable, and the range is 0 (no), 1 (yes).

Fireplace: States whether the house has a fireplace, it is a numerical variable, and the range is 0 (no), 1 (yes).

Neighborhood: What neighborhood the house is located in geographically, it is a categorical variable and there is no range.

## Step 2 Results
The variable of interest is sales price of the houses.

## Step 3 Results
```{r}
library(ggplot2)
library(dplyr)

ggplot(ames %>% filter(`Sale Price` > 0), 
       aes(x = log(get("Sale Price")))) + 
  geom_histogram(binwidth = 1)
```

## Step 4 Results

Analyn's Work:

```{r, message = FALSE}
library(classdata)
library(ggplot2)
library(dplyr)


# Scatter Plot of Sale Price vs Number of Bedrooms
ggplot(ames %>% filter(`Sale Price` > 0 & Bedrooms > 0), 
       aes(x = `Bedrooms`, 
           y = log(get("Sale Price")))) +
  geom_point(alpha = 0.5, color = "blue") +
  theme_minimal() +
  labs(title = "Scatterplot of Sale Price vs Bedrooms",
       x = "Bedrooms",
       y = "Sale Price (log)")

```

I chose to use a scatter plot to demonstrate the relationship between the Sale Price of a building and the number of Bedrooms that are located in the building. I removed the zero values for both the Sale Price and the Bedroom variables. The plot shows that the higher cost homes typically have between 2 to 5 bedrooms, which makes sense as that is the typical size of a house. The plot shows the 10 bedroom house sold for a high price, which makes sense, as bigger houses typically sell for more. There are a few outliers, where 3 and 4 bedroom homes were sold for a significantly lower price, but this can likely be attributed to the housing crash of 2008. To verify this, we would want to check the year sold of these houses.

Tirmidi's work:
## Lot Area Impact
```{r}
# Finding range of Lot Area
sum(is.na(ames$`LotArea(sf)`))

colnames(ames) # Printing all column names

# Scatterplot of Sale Price vs Lot Area
ggplot(ames, aes(x = `LotArea(sf)`, y = `Sale Price`)) +
  geom_point(alpha = 0.5, color = "blue") +
  theme_minimal() +
  labs(title = "Scatterplot of Sale Price vs Lot Area",
       x = "Lot Area (sq ft)",
       y = "Sale Price")



library(ggplot2)
# library(ggthemes) # For better themes

ggplot(ames, aes(x = `LotArea(sf)`, y = `Sale Price`)) +
  geom_hex(bins = 15) +  # Adjust bins for resolution
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(title = "Hexbin Plot: Sale Price vs Lot Area",
       x = "Lot Area (sq ft)",
       y = "Sale Price")
```

