---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

```{r, include=FALSE}
library(classdata)

```

## Step 1 Results

**What variables are there?**

Parcel ID, Address, Style, Occupancy, Sale Date, Sale Price, Multi-sale, Year Built, Acres, Total Living Area (SF), Bedrooms, Finished BSMT (SQ), Lot Area (SF), AC, Fireplace, Neighborhood

**What type are the variables?**

There is a mix of categorical (ex. Neighborhood, Occupancy, Style) and numerical (ex. Total Living Area (SF), Bedrooms, Finished BSMT (SQ), Lot Area (SF)). Some of the variables are continuous.

**What does each variable mean? What do we expect their data range to be?**

Parcel ID: The house ID number in the system, it is a categorical variable and there is no range.

Address: The address the house is located at, it is a categorical variable and there is no range.

Style: The architectural style the house was built in, it is a categorical variable, and the range is 12 (there are 12 different styles the house can be classified as)

Occupancy: How many people live in the house, it is a categorical variable, and the range is 5

Sale Date: The date the house was sold, it is a categorical variable, and the range is any date between July 3rd, 2017 and August 31st, 2022.

Sale Price: The price the building was sold for, it is a numerical variable, and the range is 0 - 22 million

Multi-sale: States if the house has been sold multiple times, it is a numerical variable, and the range is 0 (no), 1 (yes).

Year Built: The year the building was built, it is a numerical variable, and the range is 1880 to 2022 (where 0 is present as a missing value).

Acres: The total acres on the property, it is a numerical variable, and the range is 0 to 12.

Total Living Area (SF): The total living area square footage, it is a numerical variable, and the range is 0 to 6,007.

Bedrooms: The number of bedrooms in the property, it is a numerical variable, and the range is 0 to 10.

Finished BSMT (SQ): The finished basement square footage, it is a numerical variable, and the range is 10 to 6,496 (with NA for non-finished basements).

Lot Area (SF): The square footage of the lot the house was built on, it is a numerical variable, and the range is 0 to 523,228

AC: States whether the house has air conditioning, it is a numerical variable, and the range is 0 (no), 1 (yes).

Fireplace: States whether the house has a fireplace, it is a numerical variable, and the range is 0 (no), 1 (yes).

Neighborhood: What neighborhood the house is located in geographically, it is a categorical variable and there is no range.

## Step 2 Results
The variable of interest is sales price of the houses.

## Step 3 Results
```{r}
library(ggplot2)
ggplot(ames, aes(x = `Sale Price`)) + geom_histogram(binwidth = 350000)
```

## Step 4 Results

Tirmidi's work:
## Lot Area Impact
```{r}
# Finding range of Lot Area
sum(is.na(ames$`LotArea(sf)`))

colnames(ames) # Printing all column names

# Scatterplot of Sale Price vs Lot Area
ggplot(ames, aes(x = `LotArea(sf)`, y = `Sale Price`)) +
  geom_point(alpha = 0.5, color = "blue") +
  theme_minimal() +
  labs(title = "Scatterplot of Sale Price vs Lot Area",
       x = "Lot Area (sq ft)",
       y = "Sale Price")



library(ggplot2)
library(ggthemes) # For better themes

ggplot(ames, aes(x = `LotArea(sf)`, y = `Sale Price`)) +
  geom_hex(bins = 15) +  # Adjust bins for resolution
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(title = "Hexbin Plot: Sale Price vs Lot Area",
       x = "Lot Area (sq ft)",
       y = "Sale Price")
```

Ryan's Work:
#Occupancy Impact
```{r}
ggplot(ames, aes(x = Occupancy)) + geom_bar() + coord_flip()
```
Occupancy has 6 different possible values within the data. These values are 'NA', 'Condominium', Single-Family/ Owner Occupied', 'Townhouse', 'Two-Family Duplex', and 'Two-Family Conversion'. The most popular occupancy is 'Single-Family/Owner Occupied by a difference of about 4,000 more then the next two popular occupancies of 'Townhouse' and 'Condominium'.

```{r}
#ames$OccupancyNoNA <- na.omit(ames$Occupancy)
ggplot(ames, aes(x = Occupancy, weight = `Sale Price`, fill = Occupancy)) + geom_bar() + coord_flip() + labs(x = "Occupancy", y = "Sale Price", title = "Occupancy vs Sale Price")
```
'Condominium' occupancy type has the most expensive sale price by a significant amount compared to others. This is due to the fact that 'Condominium' represents the purchasing of an entire building on apartments or condominiums with many units included. The next is 'Single-Family/Owner Occupied' as these are stand-alone houses, and are the next most expensive houses. The other three occupancy types are low compared to these two, since these are typically smaller and the buildings are owned by separate individuals. The only oddity is condominium being the most expensive by a significant amount, but this is because these sales are buyers buying an entire apartment complex. 
